---
title: "Tofacitinib Pooled Analysis"
author: "Nathan Constantine-Cooke"
date: '`r Sys.Date()`'
output:
   html_document:
     code_folding: hide
     toc: true
     toc_float:
       collapsed: true
---

## Introduction

This document provides a short exploration into the feasibility of carrying out
a pooled analysis of the efficacy and safety of tofacitinib in the present day.

```{r setup, message=F}
library(meta)
library(metafor)
```

## Efficacy

Data from 10 papers was used for the analysis of the efficacy of tofacitinib. 

```{r}
efficacy <- read.csv("efficacy.csv")
knitr::kable(efficacy)
```

``rem8``, the number of patients who had entered clinical remission either 8 or
12 weeks after starting tofacitinib, was used for this exploration. To reduce
heterogeneity, only UC patients were used. Due to the small sample size, a
logit transformation was applied. 

```{r}
ies.efficacy <- escalc(xi = rem8, ni = UC, measure = "PLO", data = efficacy)
```

A random effects model was fitted using the DerSimonian and Laird method.
Testing for heterogenity gave a highly signifcant p-value (p < 0.0001).

```{r, warning=F}
efficacy.pes.logit <- rma(yi, vi, data = ies.efficacy, method = "DL")
print(efficacy.pes.logit)
```

$\tau^2$ was estimated to be 0.2860 (95% confidence interval 0.1208, 1.5796).
$I^2$ was estimated to be 0.8185 (95% confidence interval 0.6557, 0.9614). 

```{r}
confint(efficacy.pes.logit)
```

Transforming the results back to the original scale results in a prediction of
the percentage of patients who entered clinical remission at weeks 8 or 12 to be
34.82% (95% CI 26.46%, 44.22%)

```{r}
efficacy.pes <- predict(efficacy.pes.logit, transf = transf.ilogit)
print(efficacy.pes)
```


## Safety

Data from 12 papers was used for the analysis of the safety of tofacitinib.

```{r}
safety <- read.csv("safety.csv")
knitr::kable(safety)
```

``AE``, the number of patients who had an adverse event during the study, was 
used for this exploration. To reduce heterogeneity, only UC patients were used.
Due to the small sample sizes, a logit transformation was applied. 

```{r}
ies.safety <- escalc(xi = AE, ni = ni, measure = "PLO", data = safety)
```

A random effects model was fitted using the DerSimonian and Laird method.
Testing for heterogenity gave a highly signifcant p-value (p < 0.0001).

```{r}
safety.pes.logit <- rma(yi, vi, data = ies.safety, method = "DL")
print(safety.pes.logit)
```

$\tau^2$ was estimated to be 0.3165 (95% confidence interval 0.1077, 2.1381).
$I^2$ was estimated to be 0.8306 (95% confidence interval 0.6252, 0.9707). 

```{r}
confint(safety.pes.logit)
```

Transforming the results back to the original scale results in a prediction of
the proportion of patients who had an adverse event to be 21.33% (95% CI 14.60%,
30.07%)

```{r}
safety.pes <- predict(safety.pes.logit, transf = transf.ilogit)
print(safety.pes)
```